The abundance of Bullet-Groups in $\Lambda$CDM

### 2014-04-01 paper/ms.tex
new version for the co-authors
updated text
typo
submitted version

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_1_V3.eps}
\end{center}
\caption{
  Integrated probability distribution for the displacement between the
  center of the host halo and its dominant sub-halo at all redshifts
  $z=0$, $0.25$, $0.5$ and $1.0$. The left panel
  shows the results in terms of the physical displacements while the
  right panel shows the displacements normalized by the virial radius
  of the host halo. The continuous (dashed) line corresponds to the halos in the
  cluster (group) sample.
  The vertical lines show the mean value and uncertainties ($186\pm
  30\hkpc$) in the  separation between the two dark matter clumps
  estimated in \citet{Gastaldello} for the \bullg. Between $40\%$
  to $60\%$ of the groups show a displacement equal or larger than
  this observational benchmark. This fraction rises to $60\%$ and
  $80\%$ in clusters.}
\label{fig:displacement}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_3_V3.eps}
\end{center}
\caption{2D histograms in the plane $\mu$-$D_{\rm off}$. The left
  panel corresponds to clusters and the right panel to groups. The data
  used to construct this histogram includes halos at all redshifts.}
\label{fig:geometry}
\end{figure*}

@@

The results for the integrated distributions for $d_{\rm 2D}^{\rm bar}$
are shown in Figure \ref{fig:baryonic_displacements}. The dashed lines
represented the results for groups and the continuous lines correspond
to clusters. As a test of our model we compare the cluster results
against the analytic fit provided by \cite{ForeroRomero2010}. This fit
reproduces the statistics for the DM-baryon separation found for
clusters more massive than $>10^{14}$\hMsun\ in a simulation
which included a description for DM and gas with $8$ times more volume
and $8$ times less mass resolution than the Bolshoi Simulation. The
fit is valid for separations larger than $70$\hkpc, beyond which we
find that it provides a remarkably good description within a factor of
$\sim 2$ of our results. This gives us confidence in our approach to
correctly estimate the expected fraction of groups with a DM-gas
displacement.

@@

In this Letter we estimated the fraction of galaxy groups and clusters
in a $\Lambda$CDM cosmology that could present observational features
associated to a bullet-like event. This is motivated by the recent
observational results of \citep{Gastaldello} where a system (\bullg)
on the mass range $1\times 10^{14}$\hMsun\ and velocity dispersion
$650$\kms was reported to feature a displacement between its baryonic
(gas) and dark matter components.

@@

Our main result is that a fraction of $40\%$-$60\%$ of the halos
in the group sample presents DM-DM displacement equal or larger than the
observed displacement for \bullg. For halos in the cluster sample this
fraction increases to $60\%$-$80\%$. We also derived an
estimate for the DM-baryon displacement. In the group sample
$0.1\%$-$1.0\%$ of the halos show a displacement equal or larger than
the  measurements of \bullg\ by \citep{Gastaldello}; in the cluster
sample this fraction rises to $4\%$-$10\%$.

@@

Using the same criteria we find that in the simulation there there are
$\sim 6.0\times 10^{-7}$ Mpc$^{-3}$ groups similar to \bullg. This
number density is three times larger than the computed value for
clusters. This opens up a new observational possibility with surveys
such as SL2S that target a large number of groups and estimate its
multi-modal nature from lensing analysis \citep{Foex2013}. An approach
that can be further exploited with upcoming lensing surveys (e.g. with
 he {\it Euclid} satellite) and pushes for X-ray surveys with higher
sensibility: chances are larger to find bullets in systems with lower
X-ray luminosities.

@@

G.F. acknowledges support from FONDECYT through grant 3120160.

@@

\bibliographystyle{apj}
\bibliography{references}

>>>>

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_1.eps}
\end{center}
\caption{
  Integrated probability distribution for the displacement between the
  center of the host halo and its dominant sub-halo at all redshifts
  $z=0$, $0.25$, $0.5$ and $1.0$. The left panel
  shows the results in terms of the physical displacements while the
  right panel shows the displacements normalized by the virial radius
  of the host halo. The continuous (dashed) line corresponds to the halos in the
  cluster (group) sample.
  The vertical lines show the mean value and uncertainties ($186\pm
  30\hkpc$) in the  separation between the two dark matter clumps
  estimated in \citet{Gastaldello} for the \bullg. Between $40\%$
  to $60\%$ of the groups show a displacement equal or larger than
  this observational benchmark. This fraction rises to $60\%$ and
  $80\%$ in clusters.}
\label{fig:displacement}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_2.eps}
\end{center}
\caption{2D histograms in the plane $\mu$-$D_{\rm off}$. The left
  panel corresponds to clusters and the right panel to groups. The data
  used to construct this histogram includes halos at all redshifts.}
\label{fig:geometry}
\end{figure*}

@@

The results for the integrated distributions for $d_{\rm 2D}^{\rm bar}$
are shown in Figure \ref{fig:baryonic_displacements}. The dashed lines
represented the results for groups and the continuous lines correspond
to clusters. As a test of our model we compare the cluster results
against the analytic fit provided by \cite{ForeroRomero2010}. This fit
reproduces the statistics for the DM-baryon separation found for
clusters more massive than $>10^{14}$\hMsun\ in a simulation
which included a description for DM and gas with $8$ times more volume
and $8$ times less mass resolution than the Bolshoi Simulation. The
fit is valid for separations larger than $70\hkpc$, beyond which we
find that it provides a remarkably good description within a factor of
$\sim 2$ of our results. This gives us confidence in our approach to
correctly estimate the expected fraction of groups with a DM-gas
displacement.

@@

In this Letter we estimated the fraction of galaxy groups and clusters
in a $\Lambda$CDM cosmology that could present observational features
associated to a bullet-like event. This is motivated by the recent
observational results of \cite{Gastaldello} where a system (\bullg)
on the mass range $1\times 10^{14}$\hMsun\ and velocity dispersion
$650$\kms was reported to feature a displacement between its baryonic
(gas) and dark matter components.

@@

Our main result is that a fraction of $40\%$-$60\%$ of the halos
in the group sample presents DM-DM displacement equal or larger than the
observed displacement for \bullg. For halos in the cluster sample this
fraction increases to $60\%$-$80\%$. We also derived an
estimate for the DM-baryon displacement. In the group sample
$0.1\%$-$1.0\%$ of the halos show a displacement equal or larger than
the  measurements of \bullg\ by \cite{Gastaldello}; in the cluster
sample this fraction rises to $4\%$-$10\%$.

@@

Using the same criteria we find that in the simulation there are
$\sim 6.0\times 10^{-7}$ Mpc$^{-3}$ groups similar to \bullg. This
number density is three times larger than the computed value for
clusters. This opens up a new observational possibility with surveys
such as SL2S that target a large number of groups and estimate its
multi-modal nature from lensing analysis \citep{Foex2013}. An approach
that can be further exploited with upcoming lensing surveys (e.g. with
 he {\it Euclid} satellite) and pushes for X-ray surveys with higher
sensibility: chances are larger to find bullets in systems with lower
X-ray luminosities.

@@

J.G.F-T acknowledges support from Universidad de los Andes de Bogot\'a
- Colombia and Centro de Investigaciones de Astronom\'ia (CIDA) -
Venezuela.

G.F. acknowledges support from FONDECYT through grant 3120160 and
ECOS/CONICYT through grant C12U02.

@@

\bibliographystyle{apj}
%\bibliography{references} 
\begin{thebibliography}{24}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{{Brada{\v c}} {et~al.}(2008){Brada{\v c}}, {Allen}, {Treu},
  {Ebeling}, {Massey}, {Morris}, {von der Linden}, \& {Applegate}}]{Bradac2008}
{Brada{\v c}}, M., {Allen}, S.~W., {Treu}, T., {Ebeling}, H., {Massey}, R.,
  {Morris}, R.~G., {von der Linden}, A., \& {Applegate}, D. 2008, \apj, 687,
  959

\bibitem[{{Cabanac} {et~al.}(2007){Cabanac}, {Alard}, {Dantel-Fort}, {Fort},
  {Gavazzi}, {Gomez}, {Kneib}, {Le F{\`e}vre}, {Mellier}, {Pello}, {Soucail},
  {Sygnet}, \& {Valls-Gabaud}}]{Cabanac2007}
{Cabanac}, R.~A., {Alard}, C., {Dantel-Fort}, M., {Fort}, B., {Gavazzi}, R.,
  {Gomez}, P., {Kneib}, J.~P., {Le F{\`e}vre}, O., {Mellier}, Y., {Pello}, R.,
  {Soucail}, G., {Sygnet}, J.~F., \& {Valls-Gabaud}, D. 2007, \aap, 461, 813

\bibitem[{{Clowe} {et~al.}(2006){Clowe}, {Brada{\v c}}, {Gonzalez},
  {Markevitch}, {Randall}, {Jones}, \& {Zaritsky}}]{Clowe2006}
{Clowe}, D., {Brada{\v c}}, M., {Gonzalez}, A.~H., {Markevitch}, M., {Randall},
  S.~W., {Jones}, C., \& {Zaritsky}, D. 2006, \apjl, 648, L109

\bibitem[{{Dahle} {et~al.}(2013){Dahle}, {Sarazin}, {Lopez}, {Kouveliotou},
  {Patel}, {Rol}, {van der Horst}, {Fynbo}, {Wijers}, {Burrows}, {Gehrels},
  {Grupe}, {Ramirez-Ruiz}, \& {Micha{\l}owski}}]{Dahle2013}
{Dahle}, H., {Sarazin}, C.~L., {Lopez}, L.~A., {Kouveliotou}, C., {Patel},
  S.~K., {Rol}, E., {van der Horst}, A.~J., {Fynbo}, J., {Wijers}, R.~A.~M.~J.,
  {Burrows}, D.~N., {Gehrels}, N., {Grupe}, D., {Ramirez-Ruiz}, E., \&
  {Micha{\l}owski}, M.~J. 2013, \apj, 772, 23

\bibitem[{{Dawson} {et~al.}(2012){Dawson}, {Wittman}, {Jee}, {Gee}, {Hughes},
  {Tyson}, {Schmidt}, {Thorman}, {Brada{\v c}}, {Miyazaki}, {Lemaux}, {Utsumi},
  \& {Margoniner}}]{Dawson2012}
{Dawson}, W.~A., {Wittman}, D., {Jee}, M.~J., {Gee}, P., {Hughes}, J.~P.,
  {Tyson}, J.~A., {Schmidt}, S., {Thorman}, P., {Brada{\v c}}, M., {Miyazaki},
  S., {Lemaux}, B., {Utsumi}, Y., \& {Margoniner}, V.~E. 2012, \apjl, 747, L42

\bibitem[{{Farrar} \& {Rosen}(2007)}]{Farrar2007}
{Farrar}, G.~R., \& {Rosen}, R.~A. 2007, Physical Review Letters, 98, 171302

\bibitem[{{Fo{\"e}x} {et~al.}(2013){Fo{\"e}x}, {Motta}, {Limousin}, {Verdugo},
  {More}, {Cabanac}, {Gavazzi}, \& {Mu{\~n}oz}}]{Foex2013}
{Fo{\"e}x}, G., {Motta}, V., {Limousin}, M., {Verdugo}, T., {More}, A.,
  {Cabanac}, R., {Gavazzi}, R., \& {Mu{\~n}oz}, R.~P. 2013, \aap, 559, A105

\bibitem[{{Forero-Romero} {et~al.}(2010){Forero-Romero}, {Gottl{\"o}ber}, \&
  {Yepes}}]{ForeroRomero2010}
{Forero-Romero}, J.~E., {Gottl{\"o}ber}, S., \& {Yepes}, G. 2010, \apj, 725,
  598

\bibitem[{{Gastaldello} {et~al.}(2014){Gastaldello}, {Limousin}, {Foex},
  {Mu\~noz}, {Verdugo}, {Motta}, {More}, {Cabanac}, {Buote}, {Eckert},
  {Ettori}, {Fritz}, {Ghizzardi}, {Humphrey}, {Meneghetti}, \&
  M.}]{Gastaldello}
{Gastaldello}, F., {Limousin}, M., {Foex}, G., {Mu\~noz}, R., {Verdugo}, T.,
  {Motta}, V., {More}, A., {Cabanac}, R., {Buote}, D., {Eckert}, D., {Ettori},
  S., {Fritz}, A., {Ghizzardi}, S., {Humphrey}, P., {Meneghetti}, M., \& M., R.
  2014, \mnras\ accepted

\bibitem[{{Hayashi} \& {White}(2006)}]{Hayashi2006}
{Hayashi}, E., \& {White}, S.~D.~M. 2006, \mnras, 370, L38

\bibitem[{{Hinshaw} {et~al.}(2013){Hinshaw}, {Larson}, {Komatsu}, {Spergel},
  {Bennett}, {Dunkley}, {Nolta}, {Halpern}, {Hill}, {Odegard}, {Page}, {Smith},
  {Weiland}, {Gold}, {Jarosik}, {Kogut}, {Limon}, {Meyer}, {Tucker}, {Wollack},
  \& {Wright}}]{hinshaw_etal13}
{Hinshaw}, G., {Larson}, D., {Komatsu}, E., {Spergel}, D.~N., {Bennett}, C.~L.,
  {Dunkley}, J., {Nolta}, M.~R., {Halpern}, M., {Hill}, R.~S., {Odegard}, N.,
  {Page}, L., {Smith}, K.~M., {Weiland}, J.~L., {Gold}, B., {Jarosik}, N.,
  {Kogut}, A., {Limon}, M., {Meyer}, S.~S., {Tucker}, G.~S., {Wollack}, E., \&
  {Wright}, E.~L. 2013, \apjs, 208, 19

\bibitem[{{Klypin} {et~al.}(1999){Klypin}, {Gottl{\"o}ber}, {Kravtsov}, \&
  {Khokhlov}}]{BDMb}
{Klypin}, A., {Gottl{\"o}ber}, S., {Kravtsov}, A.~V., \& {Khokhlov}, A.~M.
  1999, \apj, 516, 530

\bibitem[{{Klypin} \& {Holtzman}(1997)}]{BDM}
{Klypin}, A., \& {Holtzman}, J. 1997, ArXiv Astrophysics e-prints

\bibitem[{{Klypin} {et~al.}(2011){Klypin}, {Trujillo-Gomez}, \&
  {Primack}}]{2011ApJ...740..102K}
{Klypin}, A.~A., {Trujillo-Gomez}, S., \& {Primack}, J. 2011, \apj, 740, 102

\bibitem[{{Kravtsov} {et~al.}(1997){Kravtsov}, {Klypin}, \&
  {Khokhlov}}]{Kravtsov1997}
{Kravtsov}, A.~V., {Klypin}, A.~A., \& {Khokhlov}, A.~M. 1997, \apjs, 111, 73

\bibitem[{{Lee} \& {Baldi}(2012)}]{Lee2012}
{Lee}, J., \& {Baldi}, M. 2012, \apj, 747, 45

\bibitem[{{Lee} \& {Komatsu}(2010)}]{Lee2010}
{Lee}, J., \& {Komatsu}, E. 2010, \apj, 718, 60

\bibitem[{{Markevitch} {et~al.}(2004){Markevitch}, {Gonzalez}, {Clowe},
  {Vikhlinin}, {Forman}, {Jones}, {Murray}, \& {Tucker}}]{Markevitch2004}
{Markevitch}, M., {Gonzalez}, A.~H., {Clowe}, D., {Vikhlinin}, A., {Forman},
  W., {Jones}, C., {Murray}, S., \& {Tucker}, W. 2004, \apj, 606, 819

\bibitem[{{Merten} {et~al.}(2011){Merten}, {Coe}, {Dupke}, {Massey}, {Zitrin},
  {Cypriano}, {Okabe}, {Frye}, {Braglia}, {Jim{\'e}nez-Teja}, {Ben{\'{\i}}tez},
  {Broadhurst}, {Rhodes}, {Meneghetti}, {Moustakas}, {Sodr{\'e}}, {Krick}, \&
  {Bregman}}]{Merten2011}
{Merten}, J., {Coe}, D., {Dupke}, R., {Massey}, R., {Zitrin}, A., {Cypriano},
  E.~S., {Okabe}, N., {Frye}, B., {Braglia}, F.~G., {Jim{\'e}nez-Teja}, Y.,
  {Ben{\'{\i}}tez}, N., {Broadhurst}, T., {Rhodes}, J., {Meneghetti}, M.,
  {Moustakas}, L.~A., {Sodr{\'e}}, Jr., L., {Krick}, J., \& {Bregman}, J.~N.
  2011, \mnras, 417, 333

\bibitem[{{More} {et~al.}(2012){More}, {Cabanac}, {More}, {Alard}, {Limousin},
  {Kneib}, {Gavazzi}, \& {Motta}}]{More2012}
{More}, A., {Cabanac}, R., {More}, S., {Alard}, C., {Limousin}, M., {Kneib},
  J.-P., {Gavazzi}, R., \& {Motta}, V. 2012, \apj, 749, 38

\bibitem[{{Poole} {et~al.}(2006){Poole}, {Fardal}, {Babul}, {McCarthy},
  {Quinn}, \& {Wadsley}}]{Poole2006}
{Poole}, G.~B., {Fardal}, M.~A., {Babul}, A., {McCarthy}, I.~G., {Quinn}, T.,
  \& {Wadsley}, J. 2006, \mnras, 373, 881

\bibitem[{{Riebe} {et~al.}(2013){Riebe}, {Partl}, {Enke}, {Forero-Romero},
  {Gottl{\"o}ber}, {Klypin}, {Lemson}, {Prada}, {Primack}, {Steinmetz}, \&
  {Turchaninov}}]{Riebe2013}
{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottl{\"o}ber},
  S., {Klypin}, A., {Lemson}, G., {Prada}, F., {Primack}, J.~R., {Steinmetz},
  M., \& {Turchaninov}, V. 2013, Astronomische Nachrichten, 334, 691

\bibitem[{{Springel} \& {Farrar}(2007)}]{Springel2007}
{Springel}, V., \& {Farrar}, G.~R. 2007, \mnras, 380, 911

\bibitem[{{Thompson} \& {Nagamine}(2012)}]{Thompson2012}
{Thompson}, R., \& {Nagamine}, K. 2012, \mnras, 419, 3560

\end{thebibliography}
